---
title: "PA1_template"
author: "Sofia Coradini"
date: "12/08/2020"
output: md_document
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
library(lattice)
library(Formula)
library(survival)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if (!file.exists("activity.csv") )
    {
     dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }


data <- read.csv("activity.csv")
```

-----

## What is mean total number of steps taken per day?
```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(steps_by_day$steps, xlab='Total steps per day', ylab='Frequency using binwith 500',binwidth=450, geom= "histogram") +
  theme_bw()

``` 

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMean <- mean(steps_by_day$steps)
stepsByDayMedian <- median(steps_by_day$steps)
stepsByDayMean
stepsByDayMedian

```
* Mean: 10766.19
* Median: 10765

-----

## What is the average daily activity pattern?
```{r}

steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

```
##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```
* Most Steps are 835

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
NATotal <- sum(!complete.cases(data))
NATotal
```

* Total Number of Missing values are 2304


##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
    obs <- data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}

new_activity <- data
new_activity$steps <- fillNA

```

##### 4. Make a histogram of the total number of steps taken each day 
```{r}
StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)

hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="pink", xlab="Number of Steps")


hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("pink", "red"), lwd=10)
```
##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
rmeantotal <- mean(StepsTotalUnion$steps)
rmediantotal <- median(StepsTotalUnion$steps)
rmeantotal
rmediantotal
```
* Mean (Imputed) are 10766.19
* Median (Imputed) are 10766.19

#### Do these values differ from the estimates from the first part of the assignment?
```{r}
rmeandiff <- rmeantotal - stepsByDayMean

rmediandiff <- rmediantotal - stepsByDayMedian
rmeandiff
rmediandiff
```
* The mean(Mean Var: 0) is the same, but the median does have a little variance(Median Var:1.1886792). 

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?
* The impact of the missing data has the biggest effect on the 10000 - 150000 step interval and changes frequency from 27.5 to 35, there is a variance of 7.5.

----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
```

##### 2. Make a panel plot containing a time series plot

```{r}
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```