---
title: "PA1_template"
author: "Sofia Coradini"
date: "10/08/2020"
output: html_document
---
# Loading and preprocessing the data
You can load the data with read.cvs function

```{r, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
```

And look the decriptive aspects of data with summary(), str() and head()

```{r, echo=TRUE}
summary(data)
str(data)
head(data)
```

# Process/transform the data into a format suitable for analysis - cleaning data analysis:
Converting the “date” variable to Date classe and the “interval” variable to a factor
```{r, echo=TRUE}
data$date <-as.Date(data$date, format = "%Y-%m-%d")
data$interval <-factor(data$interval)
```

# Question 01: What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
steps_by_day<-aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="thistle",xlab="Number of Steps", lwd=1)
```
Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
rmean <- mean(steps_by_day$steps)
rmean
rmedian <- median(steps_by_day$steps)
rmedian
```

# Question 02: What is the average daily activity pattern?
Steps:
1.Calculating the average
```{r, echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
```
2.Ploting the average daily activity pattern
```{r, echo=TRUE}
plot(steps_by_interval$interval,steps_by_interval$steps, type="1", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```
3. The 5-minute interval that contains the maximum number of steps
```{r, echo=TRUE}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

# Question 03: Imputing missing values
## 1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
NATotal <- sum(!complete.cases(data))
NATotal
```
## 2.Using Mean for the day compute missing values
### finding the indices of missing values (NAs)
```{r, echo=TRUE}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
```
### Imputing missing values using the mean for that 5-minute interval
```{r, echo=TRUE}
fillNA <- numeric()
for (i in 1:nrow(data)) {
  obs <- data[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(StepsAverage, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fillNA <- c(fillNA, steps)
}
```

## 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
new_activity <- data
new_activity$steps <- fillNA
```

## 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r, echo=TRUE}
StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="tomato", xlab="Number of Steps")
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="peachpuff", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("tomato", "peachpuff"), lwd=7)
```

### Calculating the mean and median 
```{r, echo=TRUE}
rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal
rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal
```
### Do these values differ from the estimates from the first part of the assignment?
```{r, echo=TRUE}
rmediandiff <- rmediantotal - rmedian
rmediandiff
rmeandiff <- rmeantotal - rmean
rmeandiff
```
### The mean is the same however the median does have a small variance.The mean and median for the **complete dataset** are almost identical.
### What is the impact of imputing missing data on the estimates of the total daily number of steps?
### The impact of the missing data has the *biggest* effect on the 10000 - 150000 step interval according by the graph comparing the **imputed** and **non-imputed** analysis.

# Question 04: There differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")

new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))

StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)

library(lattice)

xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l", bg = "white")
```
